---
title: "Shooting_Project"
author: "ZRO"
date: "2024-03-25"
output:
  pdf_document: default
  html_document: default
---
```{r load-packages, include=FALSE}
library(tidyverse)
library(formatR)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```



## Shooting Data Project
### This project involves gathering, cleaning, analyzing and plotting NYC shooting data.

Getting the data:
```{r}
download.file("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv?accessType=DOWNLOAD", destfile = "Data/shooting_data.csv")
shootings_data <- readr::read_csv("Data/shooting_data.csv")
```

Summary of the Data imported:
```{r}
summary(shootings_data)
```
The date appears to be a character type, i'll change to a proper date type.  Lon_Lat looks redundant so i'll remove that too
```{r}
shootings_data <- shootings_data %>% mutate(OCCUR_DATE = mdy(OCCUR_DATE)) %>% select(-c(Lon_Lat))
summary(shootings_data)
```

Lets look for missing data, I see there are some NAs in Jurisdiction Code, Latitude and Longitude.  I'll remove those:
```{r}
shootings_data = shootings_data[!is.na(shootings_data$"JURISDICTION_CODE") & !is.na(shootings_data$"Latitude") & !is.na(shootings_data$"Longitude"),]
summary(shootings_data)
```


## Adding some visulizations

Lets view shooting incidents over time
```{r message=FALSE, warning=FALSE}
shootings_by_year <- shootings_data %>% group_by(year(OCCUR_DATE)) %>% summarize(Total_Shootings = n())
names(shootings_by_year) <- c('Year','Total_Shootings')
ggplot(shootings_by_year, aes(x=Year,y=Total_Shootings, ymin = 0, ymax = 2500)) + geom_point()
```


This makes me wonder why there was such an uptick in shootings in 2020
  -Maybe a certain boro is most to blame?

```{r message=FALSE, warning=FALSE}
shootings_by_boro_year <- shootings_data %>% group_by(BORO,year(OCCUR_DATE), .add=TRUE) %>% summarize(Total_Shootings = n())
names(shootings_by_boro_year) <- c('Boro', 'Year', 'Total_Shootings')
shootings_by_boro_year <- shootings_by_boro_year[shootings_by_boro_year$Year == 2019 | shootings_by_boro_year$Year == 2020,]
ggplot(shootings_by_boro_year, aes(x=Year, y = Total_Shootings)) + geom_point() + facet_grid(. ~ Boro) + scale_x_continuous(limits = c(2018,2021)) + theme(axis.text.x = element_text(angle=90))
```

It seems that every borough had an increase in shootings between the two years...


## Conclusions

In this data analysis, I pulled shooting incident data from the city of New York's public repository.
  With this data, I was able to see that there was a general downward trend in shootings
  until 2020, where there was a large uptick.  To determine if shootings had risen across
  the city or were localized to one Borough, I plotted the total shooting incidents by Borough
  for those 2 years which revealed that there was an increase in all Boroughs.
  
The root cause of this trends is most likely the increased economic uncertainty 
  that went along with the pandemic.  One area of bias I may have is that I
  believe in the pandemic and its effects whereas others may not.  It would be very
  difficult to isolate this bias without significant sources of outside data and
  research.
  
Another area of bias could be derived from the fact that I do not live in a city and
  therefore can view a city as one homogeneous thing.  Isolating the city by its
  constituent parts helped me overcome that bias.

